---
title: "p8105_hw1_yw3774"
author: "Yida Wang"
date: "9/22/2021"
output: github_document
---

```{r, library}
library(tidyverse)
```

## Promble 1
Create a data frame comprised of:  
1. a random sample of size 10 from a standard Normal distribution  
2. a logical vector indicating whether elements of the sample are greater than 0  
3. a character vector of length 10  
4. a factor vector of length 10, with 3 different factor “levels”  

```{r create_df}
df_prob1 = tibble(
  samp = rnorm(10),
  vec_logical = samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "low", "low", "mid", "mid", "mid", "high", "high", "high", "high"))
  )
```

* Take the mean of each variable in my dataframe.What works and what doesn’t?

```{r mean_pull_var}
mean(pull(df_prob1, samp))
mean(pull(df_prob1, vec_logical))
mean(pull(df_prob1, vec_char))
mean(pull(df_prob1, vec_factor))
```

The mean of numbers and logical work but character or factor don't.

---
* Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

This code chunk below applies the _as.numeric_ function to the logical, character, and factor variables.

```{r try_numeric,eval=FALSE}
as.numeric(pull(df_prob1, vec_logical))
as.numeric(pull(df_prob1, vec_char))
as.numeric(pull(df_prob1, vec_factor))
```

The _as.numeric_ function turned the logical variables into type "numeric" and logical values `FALSE` and `TRUE` are turned into `0` and `1`, respectively.  
Character variables are all coerced to `NA`.  
Factor objects are turned to numbers indicating their different levels. "low"->2, "high"->1, "mid"->3.

When we taking _mean()_ function, only the logical variable and numeric variable works.  
R consider `FALSE` and `TRUE` as `0` and `1`, so this is same as the numeric variable.
The character variables can not be consider as numeric variable or coerced to numeric type by the _as.numeric_ function. So _mean()_ function didn"t work.

# Problem 2
Loads the *Penguins** dataset from the _palmerpenguins_ package.

```{r load_penguin}
data("penguins", package = "palmerpenguins")
```
* Write a short description of the penguins dataset (not the penguins_raw dataset) using inline R code, including:  
1. the data in this dataset, including names / values of important variables  
2. the size of the dataset (using nrow and ncol)  
3. the mean flipper length  

The **Penguins** dataset contains the following variables: `r ls(penguins)`.
The number of rows (observations datas) in the Penguin dataset is `r nrow(penguins)`.
The number of columns (variables) in the Penguin dataset is `r ncol(penguins)`.
Flipper Length (mm) is a `r class(pull(penguins, flipper_length_mm))` variable. The mean flipper length in mm is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)`.

This code chunk below generates a scatter plot for bill length and flipper length, each color represents a different species.

```{r scatterplot_penguin}
ggplot(penguins, aes (x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + ggtitle("Flipper Length vs. Bill Length for Penguins by Species") + theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5, lineheight = 1.5))
ggsave("scatterplot_penguin.pdf")
```